Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-04-22 paper/ms.tex
small additions to the paper

\newcommand{\documentname}{paper~}
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade a 
central topic in studies of structure formation in the Universe. The
reason is the diverse range of fields where they are helpful, LAEs can
be used as probes of reionization, tracers of large scale structure,
signposts for low metallicity stellar populations and markers of the
the galaxy formation process through cosmic history.

@@

For each model ${\mathcal M}$ we compute the LAE surface density
distributions for the three different ways of grouping the mock
fields, as described in the previous section. For each sub-volume we
project the positions of the LAE hosting halos along the $z$ direction
and calculate its surface number density in units of sources per
arcmin$^{2}$. For each number density distribution we perform a
Kolmogorov-Smirnov againsg the $12$ surface density observational
values. From this test we obtain the value $0<P<1$ to reject the null
hypothesis, namely that the two data sets come from the same
distribution. In this paper we use values of $P>0.1$ to consider that
the simulated and observed number densities come from the same
distribution.

@@

\subsection{Dark Matter Halo Number Density}
In Figure \ref{fig:laes_dist} we present the results for  the
integrated dark matter halo surface density as a function of halos
mass. Each line correspond to one of the 210 sub-volumes in the
Bolshoi simulation. The shadowed area indicates the surface density
values for LAEs allowed by the observations.

This result allows us to better understand the
expected trends for the LAEs' preferred mass and the occupation
fraction.  From this Figure we can read which models do not have a
chance reproduce the
observations. Regions in the plot where the halo surface density
values are below the observational constraint correspond to high
masses halo masses. For a LAE model ${\mathcal M}$ with a minimum mass
$M_{\rm   min}> 3 \times 10^{11}$\hMsun located in that mass range,
the surface density is too low compared with observations.

@@

The two {\tt match} and {\tt random} methods present the highest
number of matching mock surveys in the medium mass regime. However, it
is important to keep in mind that not all the mock surveys for a
sucessful model ${\mathrm M}$ present a high value $P>0.1$, only a
modest seems to be consistent with observations. This shows that the
cosmic variance is still present on the physical scales probed by
observations.

@@

The raw results of our experiments yield a prefered range for
halo masses hosting LAES is in the range $10^{10}\hMsun<M_{\rm
  min}<10^{11.5}\hMsun$, without any limitation in the  maximum mass, with an average occupation fraction as a function of the minimum mass with an approximate dependence of the kind $f_{occ}\approx 0.1 (M_{\rm min}/10^{10})^{XX}$. These results discard models that were nevertheless disfavored from the very beginning based on the results of the mass functions (see Figure 1.).

In this discussion section we consider three different ways to impose tighter constraints by making use of the results we have derived so far and using additional statistical and astrophysical constraints. The first is based on
models with high fractions of mock surveys that are consistent with
observations. The second uses observational results on the occupation
fraction. The third exploits the observational information in the Angular
Correlation Function (ACF).

\subsection{Models with the highest success rates}
\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_match_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_random_P5.pdf}
\end{center} 
\caption{ Number of models with a mininium number of mock survey
  realizations that are consistent with observations.
  \label{figure:high_success_rate}.}  
\end{figure*}

For each model there are 15 different mock survey realizations of mock. In
the loose constraints of the previous sections, each model that had at
least one (1) realization of mock survey with $P>0.05$ was considered as consistent with observations.

In Figure XX we show the number of models that have at least $n$ realizations with $P>0.05$ for the {\texttt match} and {\texttt random} methods.  This shows that there are around $550$ to $600$ different models that have at leaset one mock survey realization consistent with observations. At the other extreme, there are $80$ to $100$ models with all the 15 realizations with $P>0.05$. Here we focuse on the latter models. The best models represent $\sim 15\%$ of the number of initially considered good models.

@@

\subsection{Observational constraints on the occupation fraction}

@@

For the cosmological parameters used in this \documentname the age of
the universe between $z=3.1$ and $z=2.2$ has changed by $\sim 1$ Gyr. We assume that the physical conditions that determine
the escape fraction $f_{\rm esc}$ and the occupation fraction $f_{\rm  occ}$ remain constant over that time scale. This assumption allows us to further pick models that have an occupation fraction of $f_{\rm occ}=0.10$. Under
this selection only $18$ models can be selected. Considering an
occupation fraction $f_{\rm occ}=0.20$ another $39$ models can be
considered. This constraint help us to select $\sim 10\%$ of the original models that were considered as consistent with observations in the previous section.

@@

\subsection{Constraints from the Angular Correlation Function}

@@

\subsection{Prediction for the angular correlation function}

@@

\section{Conclusions}
In this \documentname we constrain the preferred mass for dark matter
halos hosting Lyman Alpha Emitters at a redshift $z=3.1$. We use a
method that matches the cosmic variance in the surface
density number of LAEs from differente observed fields to mock fields
based on a simplified model with three basic parameters. The halo
masses that define the halos where LAEs can be found, $M_{\rm
  min}<M_{\rm h}<M_{\rm max}$, and the fraction of the halos in this
range that are actully occupied, $f_{\rm occ}$. After a thorugh
exploration of the paremeter space we are able to constrain the mass
range of dark matter halos hosting LAEs to be in the range $<M_{\rm
  h}<$ and a corresponding occupation fraction that escales as $f_{\rm
  occ} = M_{\rm min}$.

In this letter, our estimates are based on a large volume, high
resolution, dark matter only Nbody simulation. From the halo catalogs
in the simulation a fraction of halos in a given mass range to
host LAEs. We use this information to construct mock observations to
derive the statistics that can be compared against observations. The
spatial information in this catalogs is also kept, which allows us to
compute angular correlation functions and include these results in our
analysis.

>>>>

\newcommand{\documentname}{paper~}
\newcommand{\match}{{\tt match}~}
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade a 
central topic in studies of structure formation in the Universe. The
reason is the diverse range of fields where they are helpful. LAEs can
be used as probes of reionization, tracers of large scale structure,
signposts for low metallicity stellar populations and markers of the
the galaxy formation process through cosmic history.

@@

For each model ${\mathcal M}$ we compute the LAE surface density
distributions for the three different ways of grouping the mock
fields, as described in the previous section. For each sub-volume we
project the positions of the LAE hosting halos along the $z$ direction
and calculate its surface number density in units of sources per
arcmin$^{2}$. For each number density distribution we perform a
Kolmogorov-Smirnov against the $12$ surface density observational
values. From this test we obtain the value $0<P<1$ to reject the null
hypothesis, namely that the two data sets come from the same
distribution. In this paper we use values of $P>0.1$ to consider that
the simulated and observed number densities come from the same
distribution.

@@

\subsection{Dark Matter Halo Number Density}
In Figure \ref{fig:laes_dist} we present the results for  the
integrated dark matter halo surface density as a function of halos
mass. Each line corresponds to one of the 210 sub-volumes in the
Bolshoi simulation. The shadowed area indicates the surface density
values for LAEs allowed by the observations.

This result allows us to better understand the expected trends for the
LAEs' preferred mass and the occupation fraction.  From this Figure we
can read which models do not have achance reproduce the
observations. Regions in the plot where the halo surface density
values are below the observational constraint correspond to high
masses halo masses. For a LAE model ${\mathcal M}$ with a minimum mass
$M_{\rm   min}> 3 \times 10^{11}$\hMsun located inthat mass range, the
surface density is too low compared with observations.

@@

The two {\tt match} and {\tt random} methods present the highest
number of matching mock surveys in the medium mass regime. However, it
is important to keep in mind that not all the mock surveys for a
sucessful model ${\mathrm M}$ present a high value $P>0.1$, only a
modest fraction seems to be consistent with observations. This shows that the
cosmic variance is still present on the physical scales probed by
observations. This will be considered in more detail in the discussion
seccion.

@@

\subsection{Additional Constraints}

The raw results of our experiments yield a prefered range for
halo masses hosting LAES bounded by a minimum mass $10^{10}\hMsun<M_{\rm
  min}<10^{11.5}\hMsun$ and without any limitation on the  maximum
mass. In this regime the average occupation fraction varies with an
approximate dependence of the kind $f_{occ}\approx 0.1 (M_{\rm
  min}/10^{10})^{XX}$. These results discard models that were
nevertheless disfavored from the very beginning based on the results
of the mass functions (see Figure 1.).

In this section we consider three different ways to impose
tighter constraints on this mass range by making use of the results we
have derived so far together with additional statistical and
astrophysical constraints. In the first constraint we select the
models where all the mock surveys present KS-test values consistent
with observations. The second constraints uses recent observational
results on the average occupation fraction for LAEs at high-z. The
third exploits the information in the Angular Correlation Function
(ACF).

\subsubsection{Models with the highest success rates}
\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_match_P5.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/Fig4_random_P5.pdf}
\end{center} 
\caption{ Number of models with a mininium number of mock survey
  realizations that are consistent with observations.
  \label{figure:high_success_rate}.}  
\end{figure*}

For each model there are 15 different mock survey realizations. In the
previous section we presented the models that had at least one (1)
mock survey realization with $P>0.05$.

In Figure \ref{figure:high_success_rate} we show the number of models
that have at least $n$ realizations with $P>0.05$ for the {\texttt
  match} and {\texttt random} methods.  This shows that there are
around $550$ to $600$ different models that have at leaset one mock
survey realization consistent with observations. At the other extreme,
there are $80$ to $100$ models with all the 15 realizations with
$P>0.05$. Here we focuse on the latter models. The best models
represent $\sim 15\%$ of the number of initially considered good
models.

@@

\subsubsection{Observational constraints on the occupation fraction}

@@

For the cosmological parameters used in this \documentname the age of
the universe between $z=3.1$ and $z=2.2$ has changed by $\sim 1$
Gyr. We assume that the physical conditions that determine the escape
fraction $f_{\rm esc}$ and the occupation fraction $f_{\rm  occ}$
remain constant over that time scale. This assumption allows us to
further pick models that have an occupation fraction of $f_{\rm
  occ}=0.10$. Under this selection only $18$ models can
beselected. Considering an occupation fraction $f_{\rm occ}=0.20$
another $39$ models can beconsidered. This constraint help us to
select $\sim 10\%$ of theoriginal models that were considered as
consistent with observations in the previous section.

@@

\subsubsection{Constraints from the Angular Correlation Function}

@@

@@

\subsection{Implications for galaxy formation models}

... Intrinsic emission and escape fraction.

... Star formation rate efficiency at this redshift.

... Mass dependence of the escape fraction.

\subsection{Implications for large LAEs surveys}

... The bias for the preferred halo mass.

... The scale at which cosmic variance drops.

... This can be observationally tested with HETDEX.

\subsection{On the reproducibility of our results}

... All the software to produce the results in this paper is publicly
available.

... The raw catalogs can be obtained from the MultiDark database but
can also be obtained in the repository of this paper on github.

\section{Conclusions}
In this \documentname we constrain the preferred mass for dark matter
halos hosting Lyman Alpha Emitters at a redshift $z=3.1$. We use a
method that matches the cosmic variance in the surface
density number of LAEs between mock and real observations. The mock
catalogs are based on a simplified model with three basic parameters: the halo
mass range where LAEs can be found, $M_{\rm   min}<M_{\rm h}<M_{\rm
  max}$, and the fraction of the halos in thisrange that are actully
occupied, $f_{\rm occ}$. After a thorugh exploration of the parameter
space we are able to constrain the mass range of dark matter halos
hosting LAEs to be in the range $<M_{\rm   h}<$ and a corresponding
occupation fraction that escales as $f_{\rm   occ} = M_{\rm min}$.

We use three additional constraints to reduce the allowed
range of models. The first imposes a tighter criterion to consider a
model succesful, namely that all the mock surveys for a given model
must be consistent with observations. This restriction narrows down
the allowed range of models to be.

The second constraint is based on the observational results that high
redshift LAEs have a bulk Lyman alpha escape fraction of $XX$ which
can be also interepreted as an average occupation fraction of $XX$.
